Development and Implementation of an ANN based flow law for Numerical Simulations of Thermo-mechanical Processes at High Temperatures in FEM Software

### 2022-12-14 Article.tex
Finalize first draft

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
\usepackage{gensymb}
\usepackage{amsmath}
\usepackage{accents}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{subcaption}
\DeclareRobustCommand{\w}{\mbox{\large\ensuremath{\mathsf{w}}}}
\DeclareRobustCommand{\dotp}{\boldsymbol{\cdot}}
\DeclareRobustCommand{\ccirc}{\kern0.5ex\vcenter{\hbox{$\scriptstyle\circ$}}\kern0.5ex}
\DeclareRobustCommand{\e}[1]{{\rm e}^{#1}}
\DeclareRobustCommand{\lay}[1]{^{(#1)}}
\DeclareRobustCommand{\mdot}[1]{\accentset{\mbox{\bfseries .}}{#1}}
\DeclareRobustCommand{\ie}{\emph{i.e.}\@\xspace}
\DeclareRobustCommand{\eal}{et \emph{al.}\@\xspace}
\DeclareRobustCommand{\eg}{e.g.,\@\xspace}
\DeclareRobustCommand{\RMSE}{\text{E}_\text{RMS}}
\DeclareRobustCommand{\AARE}{\text{E}_\text{AAR}}
\DeclareRobustCommand{\R}{\text{R}}
\DeclareRobustCommand{\ps}{\text{s}^{-1}}
\DeclareRobustCommand{\mr}[2]{\multirow{#1}{*}{#2}}
\DeclareRobustCommand{\MPa}{\text{MPa}}

@@

%=================================================================
% Full title of the paper (Capitalized)
\Title{Title}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Title}

@@

% Authors, for the paper (add full first names)
\Author{Olivier Pantalé *$^{,\dagger}$\orcidA{}, Firstname Lastname $^{2,\ddagger}$ and Firstname Lastname $^{2,}$*}

@@

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Olivier Pantalé, Firstname Lastname and Firstname Lastname}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Pantalé, O.; Lastname, F.; Lastname, F.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [XX] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}

% Contact information of the corresponding author
\corres{Correspondence: e-mail@e-mail.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}

% Current address and/or shared authorship
\firstnote{Current address: Affiliation 3.}
\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

@@

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{}

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}

@@

Numerical methods for simulating the behavior of structures subjected to high thermomechanical loads, as in the case of high-temperature forming of metallic materials, are generally based on the use of commercial Finite Element (FE) codes such as Abaqus or laboratory codes such as DynELA \cite{Pantale-2004-DOF}, which has been developed at LGP for many years.
These simulation softwares are based on two types of equations: conservation equations and constitutive equations.
If the first equations are generally well established on the basis of physics and mechanics, it is not the same for the second type of equations: the constitutive equations.
Thus, in a general way, the conservation equations concern the fundamental principles of physics such as the mass conservation law, the momentum law (fundamental equation) and the energy law (declined in the form of the first and second principles of thermodynamics).
By themselves, these laws are not sufficient to describe the behavior of a material or a structure subjected to thermomechanical stresses, because the nature of the material's behavior, generally translated by means of the behavior laws, is not included in the system previously proposed.
Therefore, for each type of material, it is necessary to define behavior laws whose formulation is based on observation, in order to describe the behavior of this material under external stresses.
The quality and the accuracy of the results of any numerical simulation depend on the choice of these behavior laws and on the ability of the user to identify the coefficients of these behavior laws for a given material by performing experiments under conditions close to those encountered during the real stress of the structure in service that one wishes to design.
Depending on the nature of the stresses, these tests are based on quasi-static or dynamic tensile or compression tests, tests on thermomechanical simulators such as Gleeble or impact tests using gas launchers or Hopkinson bars.

@@

The main problem that the researchers are confronted with after the phase of realization of the experimental tests concerns the choice of the flow law to use according to the observations made on these test results.
This choice of flow law is also restricted by the nature of the FE code used.
Thus, a user of the Abaqus FE code, will turn more particularly to a flow law of the Johnson-Cook \cite{Johnson-1983} type insofar as it is natively implemented in this software.
The choice of another form of flow law (Zerilli-Armstrong, or Arrhenius for example) obliges the user to program himself the computation of the flow stress $\sigma^y$ of the material by means of a VUMAT routine in FORTRAN as proposed by Liang \eal \cite{Liang-2022} for an Arrhenius type flow law.
The implementation of this VUMAT subroutine also requires the computation of the derivatives of the flow stress $\sigma^y$ with respect to the three input variables $\varepsilon^p$, $\mdot\varepsilon$ and $T$ which can quickly become relatively complex as the complexity of the flow law increases.
The choice of this flow law is therefore doubly guided by the nature of the behavior of the material on the one hand, but especially by the list of flow laws implemented natively in the FE codes.
At this time, there is not yet a flow law generic enough to cover a wide range of material behavior.

@@

From this observation, and from the necessity to select a flow law for a given type of material, then to identify the parameters of this flow law according to experimental tests, and finally to implement this flow law in the form of a user subroutine in FORTRAN in the Abaqus FE code, we have recently proposed in Pantalé \eal \cite{Pantale-2021} an alternative approach based on the ability of Artificial Neural Networks (ANNs) to behave as universal approximators.
The underlying idea is to implement a flow law described by an ANN in the form of a FORTRAN subroutine in the Abaqus FE code.
This neural network was previously trained from the data extracted from mechanical tests of the material and can directly define the value of the flow stress $\sigma^y$ as a function of the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$.
After this learning phase based on the use of the Python library Tensorflow \cite{Abadi-2016}, the weights and biases of the neural network are transcoded into a subroutine in FORTRAN which is compiled and linked with the libraries of the Abaqus FE code in order to include the behavior of the material by allowing the computation of the flow stress $\sigma^y$ as a function of $\varepsilon^p$, $\mdot\varepsilon$ and $T$, and of its three derivatives $\partial\sigma^y/\partial\varepsilon^p$, $\partial\sigma^y/\partial\mdot\varepsilon$ and $\partial\sigma^y/\partial T$.

The structure of this paper is as follows.
Section \ref{sec:Training} deals with the presentation of a Neural Network based flow law and its training from the data reported in figure \ref{fig:Zhou-OriginalData}.
The comparison of several neural network architectures regarding accuracy and implementation complexity will be presented.
In Section \ref{sec:Implementation}, we will present the conversion of this neural network into FORTRAN and C++ subroutines for Abaqus and DynELA FE codes.
Section \ref{sec:Simulations} will present real test cases of the use of an ANN flow law in numerical simulations and the comparison between the implementations made on Abaqus and DynELA.
Finally, a conclusion and perspective section will be presented at the end.

@@

In this section, we briefly recall, as an introduction, some basic principles of Artificial Neural Networks that are relevant to this work.
The global architecture chosen to model the behavior of a material is based on a multi-layer feed-forward ANN which as proposed by Hornik \eal \cite{Hornik-1989-MFN} can be used as a universal approximator.
The architecture retained for this study concerns a neural network with two hidden layers containing a variable number of neurons on these two layers, $3$ inputs corresponding to the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$ respectively and a single output concerning the flow stress of the material $\sigma^y$.
Figure \ref{fig:ANN-scheme} shows a graphical representation of the global architecture of this neural network.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.7\columnwidth]{Figures/ANN-scheme-2HL}
\caption{Global structure of the ANN flow law with two hidden layers, 3 inputs and one output.}
\label{fig:ANN-scheme}
\end{figure}
The choice of the number of neurons in the two hidden layers is free, but must be reasonable.
Indeed, the more neurons the network contains, the more it will be able to reproduce faithfully the training data, but the less it will be able to generalize to new data (the classical problem of overlearning of neural networks).
Moreover, the more neurons it contains, the more complex its mathematical structure will be, and the more computation time it will require within the routine included in the FE code.
It is therefore necessary to respect a balance between the capacity of the network to minimize errors during the learning phase, its complexity and the speed of execution once it is transcribed into the FE code.

@@

According to Figure \ref{fig:ANN-scheme}, the neural network has $3$ inputs (referred as the input vector $\overrightarrow{x}$ of the ANN) corresponding to the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$ respectively.
These inputs are first normalized within the range $[0,1]$ in order to avoid a ill-conditioning of the system as presented by many other authors in the literature \cite{Lin-2008-ANN, Lu-2011-ANN} since these three variables represent different physical data with a very different amplitudes ($0.7$ for the plastic strain, $100$ for the strain rate and $550~\celsius$ for the temperature in the case of the data reported in Figure \ref{fig:Zhou-OriginalData}).
Therefore, the three components of the input vector $\overrightarrow{x}$ are obtained from the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$ using the following expressions:
\begin{equation}
\overrightarrow{x}=
\begin{cases}
x_1 = \frac{\varepsilon^p - [\varepsilon^p]_{min}}{[\varepsilon^p]_{max} - [\varepsilon^p]_{min}}\\
x_2 = \frac{\ln(\mdot\varepsilon/\mdot\varepsilon_0)-[\ln(\mdot\varepsilon/\mdot\varepsilon_0)]_{min}}{[\ln(\mdot\varepsilon/\mdot\varepsilon_0)]_{max}-[\ln(\mdot\varepsilon/\mdot\varepsilon_0)]_{min}}\label{eq:CR1}\\
x_3 = \frac{T-[T]_{min}}{[T]_{max}-[T]_{min}}
\end{cases}
\end{equation}
where $[~]_{min}$ and $[~]_{max}$  are the boundaries of the range of the corresponding field.
Concerning the strain rate $\mdot\varepsilon$, and taking into account that its amplitude in a real case can reach $10^5~\ps$, as proposed in Pantalé \cite{Pantale-2021}, we chose at first to substitute $\ln(\mdot\varepsilon/\mdot\varepsilon_0)$, with $\mdot\varepsilon_0$ equal to the lowest strain rate test, for the value of $\mdot\varepsilon$.
After normalization, these three input variables are introduced into the neural network and are propagated within it by the feed-forward propagation mechanism.

@@

As introduced in Section \ref{sec:Introduction}, the implementation of a flow law in a FE code requires both the computation of the flow stress $\sigma^y$ as a function of the input data, but also the evaluation of the three derivatives of $\sigma^y$ with respect to the input data in order to be able to use a Newton-Raphson algorithm within the stress integration scheme, as proposed by Ming \eal \cite{Ming-2018-ERV, Liang-2022} based on the radial return algorithm in the Abaqus or DynELA FE codes.

@@

\frac{\partial \sigma(x)}{\partial x} = \frac{\sigma(x+\delta x) - \sigma(x)}{\delta x}

@@

Given equations \ref{eq:CR1} to \ref{eq:CR2}, we can then establish in the case of a neural network containing two hidden layers and a sigmoid activation function on the two hidden layers that the derivative of $\sigma^y$ with respect to the input data $\varepsilon^p$, $\mdot\varepsilon$ and $T$ is given by the following equations. First we compute the internal terms of the ANN to compute the derivative of the ANN with respect to the input vector $\overrightarrow{x}$:

@@

@@

In the context of neural network learning, it is necessary to define the objective function to be minimized and the evaluation of the model error. In this study, the error evaluation is based on the Root Mean Square Error ($\RMSE$) given by the following equation:
\begin{equation}
\RMSE (\MPa) = \sqrt{\frac{1}{N} \sum_{i=1}^{N} \left(\square_i^e - \square_i^y\right)^2}, \label{eq:RMSE}
\end{equation}
where $N$ is the total number of numerical training data used, $\square^y$ is the value predicted by the neural network, and $\square^e$ is the corresponding experimental value of coming from the experimental tests.
The accuracy and predictive ability of the models is assessed by the Mean Absolute Relative Error ($\AARE$) defined by Equation (\ref{eq:AARE}):
\begin{equation}
\AARE(\%) = \frac{1}{N} \sum_{i=1}^{N}{\left|\frac{\square_i^y -\square_i^e}{\square_i^e}\right|} \times 100, \label{eq:AARE}
\end{equation}

@@

Once the identification phase has been completed, it is now necessary to transpose this ANN model into a subroutine in FORTRAN or C++ that can be used by a FE code such as Abaqus or DynELA. 
This is the subject of the next section.

%------------------------------------------------------------------------------------
\section{Implementation of the ANN flow law in FE softwares}\label{sec:Implementation}
%------------------------------------------------------------------------------------

\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\columnwidth]{Figures/StressUpdate}
\caption{General flowchart of a FE code, focus on the stresses computation.}
\label{fig:StressUpdate}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=0.35\columnwidth]{Figures/RadialReturn}
\caption{The Radial return algorithm used to restore consistency of the stress tensor in FE simulation.}
\label{fig:RadialReturn}
\end{figure}

%------------------------------------------------------------------------------------
\section{Numerical simulations and benchmarks tests}\label{sec:Simulations}
%------------------------------------------------------------------------------------

>>>>

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
\usepackage{gensymb}
\usepackage{amsmath}
\usepackage{accents}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{subcaption}
\DeclareRobustCommand{\w}{\mbox{\large\ensuremath{\mathsf{w}}}}
\DeclareRobustCommand{\A}{\mbox{\large\ensuremath{\mathsf{A}}}}
\DeclareRobustCommand{\B}{\mbox{\large\ensuremath{\mathsf{B}}}}
\DeclareRobustCommand{\C}{\mbox{\large\ensuremath{\mathsf{C}}}}
\DeclareRobustCommand{\Dev}{\mbox{\Large\ensuremath{\mathsf{s}}}}
\DeclareRobustCommand{\Sig}{\mbox{\Large\ensuremath{\sigma}}}
\DeclareRobustCommand{\dotp}{\boldsymbol{\cdot}}
\DeclareRobustCommand{\ccirc}{\kern0.5ex\vcenter{\hbox{$\scriptstyle\circ$}}\kern0.5ex}
\DeclareRobustCommand{\e}[1]{{\rm e}^{#1}}
\DeclareRobustCommand{\lay}[1]{^{(#1)}}
\DeclareRobustCommand{\mdot}[1]{\accentset{\mbox{\bfseries .}}{#1}}
\DeclareRobustCommand{\ie}{\emph{i.e.}\@\xspace}
\DeclareRobustCommand{\eal}{et \emph{al.}\@\xspace}
\DeclareRobustCommand{\eg}{e.g.,\@\xspace}
\DeclareRobustCommand{\RMSE}{\text{E}_\text{RMS}}
\DeclareRobustCommand{\AARE}{\text{E}_\text{AAR}}
\DeclareRobustCommand{\R}{\text{R}}
\DeclareRobustCommand{\ps}{\text{s}^{-1}}
\DeclareRobustCommand{\mr}[2]{\multirow{#1}{*}{#2}}
\DeclareRobustCommand{\MPa}{\text{MPa}}

@@

%=================================================================
% Full title of the paper (Capitalized)
\Title{Development and Implementation of an ANN based flow law for Numerical Simulations of Thermo-mechanical Processes at High Temperatures in FEM Software}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Development and Implementation of an ANN based flow law for Numerical Simulations Thermo-mechanical Processes at High Temperatures in FEM Software}

@@

% Authors, for the paper (add full first names)
\Author{Olivier Pantalé \orcidA{}}

@@

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Olivier Pantalé}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Pantalé, O.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [XX] after \address if there is only one affiliation.)
\address[1]{Laboratoire Génie de Production, Institut National Polytechnique/Ecole Nationale d'Ingénieurs de Tarbes, Université de Toulouse, 47 Av d'Azereix, F-65016, Tarbes, France}

% Contact information of the corresponding author
\corres{Correspondence: Olivier.Pantale@enit.fr; Tel.: +33-562442933}

% Current address and/or shared authorship

% The commands \thirdnote{} till \eighthnote{} are available for further notes

@@

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{Numerical methods based on Finite Element (FE) have proven their efficiency for many years in the thermomechanical simulation of forming processes.
Nevertheless, the application of these methods to new materials requires the identification and implementation of constitutive and flow laws within FE codes which sometimes pose problems, in particular because of the very strongly non-linear character of the behavior of these materials.
Computational techniques based on Machine Learning and Artificial Neural Networks are becoming more and more important in the development of these models and help the FE codes to integrate more complex behavior.
In this paper, we present the development, implementation and use of an Artificial Neural Network (ANN) based flow law for a GrC15 alloy under high temperature thermomechanical sollicitations.
The flow law modeling by ANN shows a significant superiority in terms of model prediction quality compared to classical approaches based on widely used Johnson-Cook or Arrhenius models.
Once the ANN parameters have been identified on the base of experiments, the implementation of this flow law in a Finite Element code shows promising results in terms of solution quality and respect of the material behavior.}

% Keywords
\keyword{ANN flow law; Constitutive behavior; Radial Return algorithm; Numerical implementation; VUHARD ; GrC15; Abaqus Explicit}

@@

Numerical methods for simulating the behavior of structures subjected to high thermomechanical loads, as in the case of high-temperature forming of metallic materials, are generally based on the use of commercial Finite Element (FE) codes such as Abaqus or laboratory codes such as DynELA \cite{Pantale-2004}, which has been developed in our Laboratory for many years.
These simulation codes are based on two types of equations: conservation equations and constitutive equations.
If the first equations are generally well established on the basis of physics and mechanics, it is not the same for the second type of equations: the constitutive equations.
Thus, in a general way, the conservation equations concern the fundamental principles of physics such as the mass conservation law, the momentum law (fundamental equation) and the energy law (declined in the form of the first and second principles of thermodynamics).
By themselves, these laws are not sufficient to describe the behavior of a material or a structure subjected to thermomechanical stresses, because the nature of the material's behavior, generally translated by means of the behavior laws, is not included in the system previously proposed.
Therefore, for each type of material, it is necessary to define behavior laws whose formulation is based on observation, in order to describe the behavior of this material under external stresses.
The quality and the accuracy of the results of any numerical simulation depend on the choice of these behavior laws and on the ability of the user to identify the coefficients of these behavior laws for a given material by performing experiments under conditions close to those encountered during the real stress of the structure in service that one wishes to design.
Depending on the nature of the stresses, these tests are based on quasi-static or dynamic tensile or compression tests, tests on thermomechanical simulators such as Gleeble or impact tests using gas launchers or Hopkinson bars.

@@

The main problem that the researchers are confronted with after the phase of realization of the experimental tests concerns the choice of the flow law to use according to the observations made on these test results.
This choice of flow law is also restricted by the nature of the FE code used.
Thus, a user of the Abaqus FE code, will turn more particularly to a flow law of the Johnson-Cook \cite{Johnson-1983} type insofar as it is natively implemented in this software.
The choice of another form of flow law (Zerilli-Armstrong, or Arrhenius for example) obliges the user to program himself the computation of the flow stress $\sigma^y$ of the material by means of a VUMAT routine in FORTRAN as proposed by Gao \eal \cite{Gao-2007-FRT}, Ming \eal \cite{Ming-2018} for a Johnson-Cook flow law or Liang \eal \cite{Liang-2022} for an Arrhenius type flow law.
The implementation of this VUMAT subroutine also requires the computation of the derivatives of the flow stress $\sigma^y$ with respect to the three input variables $\varepsilon^p$, $\mdot\varepsilon$ and $T$ which can quickly become relatively complex as the complexity of the flow law increases.
The choice of this flow law is therefore doubly guided by the nature of the behavior of the material on the one hand, but especially by the list of flow laws implemented natively in the FE codes.
At this time, there is not yet a flow law generic enough to cover a wide range of material behavior.

@@

From this observation, and from the necessity to select a flow law for a given type of material, then to identify the parameters of this flow law according to experimental tests, and finally to implement this flow law in the form of a user subroutine in FORTRAN in the Abaqus FE code, we have recently proposed in Pantalé \eal \cite{Pantale-2021} an alternative approach based on the ability of Artificial Neural Networks (ANNs) to behave as universal approximators as reported by Hornik \eal \cite{Hornik-1989}.
The underlying idea is to implement a flow law described by an ANN in the form of a FORTRAN subroutine in the Abaqus FE code.
This neural network was previously trained from the data extracted from mechanical tests of the material and can directly define the value of the flow stress $\sigma^y$ as a function of the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$.
After this learning phase based on the use of the Python library Tensorflow \cite{Abadi-2016}, the weights and biases of the neural network are transcoded into a subroutine in FORTRAN which is compiled and linked with the libraries of the Abaqus FE code in order to include the behavior of the material by allowing the computation of the flow stress $\sigma^y$ as a function of $\varepsilon^p$, $\mdot\varepsilon$ and $T$, and of its three derivatives $\partial\sigma^y/\partial\varepsilon^p$, $\partial\sigma^y/\partial\mdot\varepsilon$ and $\partial\sigma^y/\partial T$.

The structure of this paper is as follows.
Section \ref{sec:Training} deals with the presentation of a Neural Network based flow law and its training from the data reported in figure \ref{fig:Zhou-OriginalData}.
The comparison of several neural network architectures regarding accuracy and implementation complexity will be presented.
In Section \ref{sec:Use}, we will present the conversion of this neural network into a FORTRAN or subroutine for the Abaqus FE code.
Validation is based on real test cases of the use of an ANN flow law in numerical simulations and the comparison between the implementations made on Abaqus Explicit.
Finally, a conclusion and perspective section will be presented at the end.

@@

In this section, we briefly recall, as an introduction, some basic principles of Artificial Neural Networks that are relevant to this work.
The global architecture chosen to model the behavior of a material is based on a multi-layer feed-forward ANN which as proposed by Hornik \eal \cite{Hornik-1989} can be used as a universal approximator.
The architecture retained for this study concerns a neural network with two hidden layers containing a variable number of neurons on these two layers, $3$ inputs corresponding to the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$ respectively and a single output concerning the flow stress of the material $\sigma^y$.
Figure \ref{fig:ANN-scheme} shows a graphical representation of the global architecture of this neural network.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.7\columnwidth]{Figures/ANN-scheme-2HL}
\caption{Global structure of the ANN flow law with two hidden layers, 3 inputs and one output.}
\label{fig:ANN-scheme}
\end{figure}
The choice of the number of neurons in the two hidden layers is free, but must be reasonable.
Indeed, the more neurons the network contains, the more it will be able to reproduce faithfully the training data, but the less it will be able to generalize to new data (the classical problem of overlearning of neural networks).
Moreover, the more neurons it contains, the more complex its mathematical structure will be, and the more computation time it will require within the routine included in the FE code.
It is therefore necessary to respect a balance between the capacity of the network to minimize errors during the learning phase, its complexity and the speed of execution once it is transcribed into the FE code.

@@

According to Figure \ref{fig:ANN-scheme}, the neural network has $3$ inputs (referred as the input vector $\overrightarrow{x}$ of the ANN) corresponding to the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$ respectively.
These inputs are first normalized within the range $[0,1]$ in order to avoid a ill-conditioning of the system as presented by many other authors in the literature \cite{Lin-2008, Lu-2011} since these three variables represent different physical data with a very different amplitudes ($0.7$ for the plastic strain, $100$ for the strain rate and $550~\celsius$ for the temperature in the case of the data reported in Figure \ref{fig:Zhou-OriginalData}).
Therefore, the three components of the input vector $\overrightarrow{x}$ are obtained from the plastic strain $\varepsilon^p$, the strain rate $\mdot\varepsilon$ and the temperature $T$ using the following expressions:
\begin{equation}
\overrightarrow{x}=
\begin{cases}
x_1 = \frac{\varepsilon^p - [\varepsilon^p]_{min}}{[\varepsilon^p]_{max} - [\varepsilon^p]_{min}}\\
x_2 = \frac{\ln(\mdot\varepsilon/\mdot\varepsilon_0)-[\ln(\mdot\varepsilon/\mdot\varepsilon_0)]_{min}}{[\ln(\mdot\varepsilon/\mdot\varepsilon_0)]_{max}-[\ln(\mdot\varepsilon/\mdot\varepsilon_0)]_{min}}\\
x_3 = \frac{T-[T]_{min}}{[T]_{max}-[T]_{min}}
\end{cases},
\label{eq:CR1}
\end{equation}
where $[~]_{min}$ and $[~]_{max}$  are the boundaries of the range of the corresponding field.
Concerning the strain rate $\mdot\varepsilon$, and taking into account that its amplitude in a real case can reach $10^5~\ps$, as proposed in Pantalé \cite{Pantale-2021}, we chose at first to substitute $\ln(\mdot\varepsilon/\mdot\varepsilon_0)$, with $\mdot\varepsilon_0$ equal to the lowest strain rate test, for the value of $\mdot\varepsilon$.
After normalization, these three input variables are introduced into the neural network and are propagated within it by the feed-forward propagation mechanism.

@@

As introduced in Section \ref{sec:Introduction}, the implementation of a flow law in a FE code requires both the computation of the flow stress $\sigma^y$ as a function of the input data, but also the evaluation of the three derivatives of $\sigma^y$ with respect to the input data in order to be able to use a Newton-Raphson algorithm within the stress integration scheme, as proposed by many authors \cite{Ponthot-2002, Ming-2018, Liang-2022, Simo-1998} based on the Radial-Return algorithm in the Abaqus FE code.

@@

\frac{\partial \sigma(x)}{\partial x} = \frac{\sigma(x+\delta x) - \sigma(x)}{\delta x},

@@

Given equations \ref{eq:CR1} to \ref{eq:CR2}, we can then establish in the case of a neural network containing two hidden layers and a sigmoid activation function on the two hidden layers that the derivative of $\sigma^y$ with respect to the input data $\varepsilon^p$, $\mdot\varepsilon$ and $T$ is given by the following equations.
First we compute the internal terms of the ANN to compute the derivative of the ANN with respect to the input vector $\overrightarrow{x}$:

@@

where $\ccirc$ is the element-wise product, known as the Hadamard product, which is a binary operation that takes two matrices $\A$ and $\B$ of the same dimensions and produces another matrix $\C$ of the same dimension as the operands, where each element $C_i=A_i~B_i$.

@@

In the context of neural network learning, it is necessary to define the objective function to be minimized and the evaluation of the model error.
In this study, the error evaluation is based on the Root Mean Square Error ($\RMSE$) given by the following equation:
\begin{equation}
\RMSE (\MPa) = \sqrt{\frac{1}{N} \sum_{i=1}^{N} \left(\square_i^e - \square_i^y\right)^2}, \label{eq:RMSE}
\end{equation}
where $N$ is the total number of numerical training data used, $\square^y$ is the value predicted by the neural network, and $\square^e$ is the corresponding experimental value of coming from the experimental tests.
The accuracy and predictive ability of the models is assessed by the Mean Absolute Relative Error ($\AARE$) defined by Equation (\ref{eq:AARE}):
\begin{equation}
\AARE(\%) = \frac{1}{N} \sum_{i=1}^{N}{\left|\frac{\square_i^y -\square_i^e}{\square_i^e}\right|} \times 100 \label{eq:AARE}
\end{equation}

@@

Once the identification phase has been completed, it is now necessary to transpose this ANN model into a subroutine in FORTRAN or C++ that can be used by a FE code such as Abaqus (for the FORTRAN version) or DynELA (for the C++ version).
This is the subject of the next section.

%------------------------------------------------------------------------------------
\section{ANN flow law implementation in FE software}\label{sec:Use}
%------------------------------------------------------------------------------------

Once the neural network has been trained as presented in Section \ref{sec:ANN-traning}, it can be used in a Finite Element code for the numerical simulation of a structure subjected to thermomechanical loading.
This sentence requires the extraction of the internal variables of the neural network and their transfer in the form of a subroutine in FORTRAN or in C++ based on equations proposed in Sections \ref{sec:ANN-equations} and \ref{sec:ANN-derivative}.

%------------------------------------------------------------------------------------
\subsection{Implementation of the ANN flow law}\label{sec:Implementation}
%------------------------------------------------------------------------------------

If we refer to the general flowchart of a finite element code as shown on Figure \ref{fig:StressUpdate}, the integration of the flow law described by the ANN concerns the computation of the stresses, in the yellow rectangle on the figure.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.9\columnwidth]{Figures/StressUpdate}
\caption{General flowchart of a FE code, focus on the stresses computation.}
\label{fig:StressUpdate}
\end{figure}
Within the framework of a FE formulation in large deformations such as classically used in thermomechanical modeling of processes, this computation of the stress tensor is to be carried out on all the integration points of each element of the studied structure.
Since the numerical model can include thousands of elements, themselves comprising between $1$ and $8$ integration points depending on the nature of the elements used, this stress computation must be as fast as possible in order not to increase the CPU time too much, but precise enough so that the results are in accordance with the physics of the process.
This is even more important if we want to integrate this flow law in an explicit FE code such as Abaqus/Explicit for which one second of physical simulation corresponds to several million iterations of these stress computations.
Thus, the complexity of the ANN, \ie the number of computational steps that must be performed to compute the flow stress as a function of the input variables, is a major parameter in the choice of the neural network.
As an example, for the model presented in section \ref{sec:ANN-traning}, $180$ internal variables, $15$ neurons on the first hidden layer and $7$ neurons on the second hidden layer were listed.
Given the equations described in Sections \ref{sec:ANN-equations} and \ref{sec:ANN-derivative}, it will be necessary to compute $22$ exponentials, to make matrix-vector products of size $15\times3$ and $15\times7$ plus many other numerical operations in order to compute the flow stress $\sigma^y$ and the $3$ derivatives of it with respect to $\varepsilon^p$, $\mdot\varepsilon$ and $T$.

The implementation of the ANN flow law identified above is realized here in a VUHARD subroutine used by the Abaqus Explicit FE code in order to allow a user to program the computation of the flow stress $\sigma^y$ and its 3 derivatives as a function of the model input data.
This subroutine is used when calculating the stress tensor $\Sig_1$ at the end of an increment from the stress tensor at the beginning of the increment $\Sig_0$, the deformations, the material parameters and the history of the deformation at each finite element integration points, according to the stress integration algorithm based on the Radial-Return method as described in Simo \eal Ming \eal \cite{Ming-2018} for Abaqus explicit or Pantalé \eal \cite{Pantale-2004} for the DynELA FE code.
Thus, without going into too much detail about the stress integration scheme used in finite element codes (the curious reader can refer to \cite{Ponthot-2002, Ming-2018, Pantale-2004, Liang-2022} for details about this method), Figure \ref{fig:RadialReturn} shows the location of the VUHARD subroutine used to compute the flow stress $\sigma^y$ and its derivative $\partial\sigma^y/\partial\Gamma$ used in the writing of the two quantities $\gamma(\Gamma)$ and $\gamma^{'}(\Gamma)$ from the following relation:
\begin{equation}
\frac{d\sigma^{y}}{d\Gamma} = \sqrt{\frac{2}{3}}\left(\frac{\partial\sigma^{y}}{\partial\varepsilon^{p}}+\frac{1}{\Delta t}\frac{\partial\sigma^{y}}{\partial\mdot{\varepsilon}^p}+\frac{\eta\sigma^{y}}{\rho C_{p}}\frac{\partial\sigma^{y}}{\partial T}\right)\label{eq:derivatives}
\end{equation}
where $\Gamma$ is the consistency parameter used in the Radial-Return algorithm as defined by Simo \eal \cite{Simo-1998}, $\Delta t$ is the time increment, $\eta$ is the Taylor-Quinney coefficient defining the amount of plastic work converted into heat energy, $C_{p}$ is the specific heat coefficient, $\rho$ is the density of the material and $\partial \sigma^y/\partial\varepsilon^p$, $\partial \sigma^y/\partial\mdot\varepsilon$ and $\partial \sigma^y/\partial T$ are the three derivatives defined in Equation (\ref{eq:DANN3}).
\begin{figure}[!ht]
\centering
\includegraphics[width=0.6\columnwidth]{Figures/VumatGeneralNewtonRaphson}
\caption{General flowchart of the Radial-Return algorithm to compute the final stress tensor $\Sig_1$.}
\label{fig:RadialReturn}
\end{figure}

%------------------------------------------------------------------------------------
\subsection{Numerical simulations and benchmarks tests}\label{sec:Simulations}
%------------------------------------------------------------------------------------

In order to validate the proposed approach and to compare the different neural network architectures proposed in section \ref{sec:ANN-traning}, we propose here to model on the Abaqus/Explicit FE code the high-temperature compression of a cylinder on a Gleeble-type thermomechanical device.
We consider a cylinder in compression with an initial diameter $d_0=8~\text{mm}$ and length $h_0=12~\text{mm}$ made of GCr15 material, for which the final height after compression is $h=6~\text{mm}$, that is a reduction of $50\%$ of its total height.
The compression of the sample is done in $10$ seconds so that the strain rate  $\mdot\varepsilon$ is in the corresponding range of the characterization of the material behavior.
Concerning the flow laws, the 4 models presented in Section \ref{sec:ANN-traning} will be used and compared between them.
It is not possible here to compare these results with experimental results, as these are not available in the references of the work of Zhou \eal \cite{Zhou-2019} or Ji \eal \cite{Ji-2018}.

The mesh of the sample is made with $850$ axisymmetric quadrilateral finite elements with $4$ nodes and reduced integration (CAX4R) with $50$ elements in the vertical direction and $17$ elements in the radial direction respectively.
The cylinder is between two rigid surfaces and a Coulomb friction law with a friction coefficient at the contact surfaces was set to $\mu=0.15$.
The simulation time being fixed at $10~\text{s}$, in order to reduce the simulation time, considering that an explicit integration scheme is used, a global mass scaling is used for all simulations.
